---
title: "05Project_01"
author: "Adolfo Hoyos Díaz"
date: "May 4, 2017"
output: html_document
---

# Dependencies

```{r echo=FALSE}
source("~/code/fito_lib/r/pkgtest.R")
pkgTest('dendextend')
pkgTest('colorspace')
pkgTest('knitr')
pkgTest('ggplot2')
library(dendextend)
library(colorspace)
library(knitr)
```

# Load data

We start by creating a variable to store the number of cluster on which we are basing our clustering analysis

```{r}
k <- 4
```

Now, we load the provided files:

```{r}


pathways.values <- as.data.frame(read.table("../data/pathways.values"))
pathways.scores <- as.data.frame(read.table("../data/pathways.scores"))


```

And show some information about the **scores** file, which is a PCA version of the **pathways.values** file:

```{r}
summary(pathways.scores)
colnames(pathways.scores)
```




# Hierarchical clustering

In this section we will perform a hierarchical clustering of both datasets. For that, we define a function that does this using a complete hierarchical scheme, and Euclidean distance:

```{r}
my.hclust <- function(x, method = "complete", dmeth = "euclidean"){
  hclust(dist(x, method = dmeth), method = method)
}
```


And, of course, we apply this function to both datasets:

```{r}
hierarch.clusters.reduced <- my.hclust(pathways.scores)
hierarch.clusters.structural <- my.hclust(pathways.values)
```

The name "*reduced*" refers to the reduced nature of the **scores** dataset, which was generated using PCA over the **pathways.values** file. The latter is an original dataset containing structural information.

Now, we transform these datasets to dendogram representations and add some color

```{r}
hierarch.cluster.dendrogram.reduced <- as.dendrogram(hierarch.clusters.reduced)
hierarch.cluster.dendrogram.reduced <- color_branches(hierarch.cluster.dendrogram.reduced, k = k)
plot(hierarch.cluster.dendrogram.reduced, main = "Hierarchical clustering, reduced data")

hierarch.cluster.dendrogram.structural <- as.dendrogram(hierarch.clusters.structural)
hierarch.cluster.dendrogram.structural <- color_branches(hierarch.cluster.dendrogram.structural, k = k)
plot(hierarch.cluster.dendrogram.structural, main = "Hierarchical clustering, structural data")

```

As this clusering method produces as much branches as datapoints, we cut these trees to only have *k* branches:

```{r}
hierarch.clusters.cut.reduced <- cutree(hierarch.cluster.dendrogram.reduced, k)
hierarch.clusters.cut.structural <- cutree(hierarch.cluster.dendrogram.structural, k)
```

And we visualize both results

```{r}
ggplot(pathways.values, aes(NC, CO, color = as.factor(hierarch.clusters.cut.structural))
) + geom_point() + ggtitle("Structural data: Hierarchical clustering")

ggplot(pathways.scores, aes(RC1, RC2, color = as.factor(hierarch.clusters.cut.reduced))
) + geom_point() + ggtitle("Reduced data: Hierarchical clustering")
```

Finally, we visualize how many datapoints were grouped in the clusters:

```{r}
plot.colors <- rainbow_hcl(k)

hierarch.clusters.table.structural <- table(hierarch.clusters.cut.structural)
bp <- barplot(hierarch.clusters.table.structural, col = plot.colors,
              xlab = "clusters", ylab = "number of elements", 
              main = "structural data: hierarchical clustering histogram")
text(bp, 0, hierarch.clusters.table.structural, cex = 1, pos = 3)

hierarch.clusters.table.reduced <- table(hierarch.clusters.cut.reduced)
bp <- barplot(hierarch.clusters.table.reduced, col = plot.colors, 
              xlab = "clusters", ylab = "number of elements", 
              main = "reduced data: hierarchical clustering histogram")
text(bp, 0, hierarch.clusters.table.reduced, cex = 1, pos = 3)
```

# K-means clustering

```{r}

kmeans.clusters.structural <- kmeans(pathways.values, centers = 4, nstart = 30)
kmeans.clusters.structural$cluster <- as.factor(kmeans.clusters.structural$cluster)

ggplot(pathways.values, aes(NC, CO, color = kmeans.clusters.structural$cluster)
       ) + geom_point() + ggtitle("Structural data: K-means clustering")

kmeans.clusters.reduced <- kmeans(pathways.scores, centers = 4, nstart = 30)
kmeans.clusters.reduced$cluster <- as.factor(kmeans.clusters.reduced$cluster)

ggplot(pathways.scores, aes(RC1, RC2, color = kmeans.clusters.reduced$cluster)
) + geom_point() + ggtitle("Reduced data: K-means clustering")

```



```{r}
kmeans.clusters.table.structural <- table(kmeans.clusters.structural$cluster)

bp <- barplot(kmeans.clusters.table.structural, col = plot.colors, xlab = "clusters", ylab = "number of elements", 
              main = "Structural data: k-means histrogram")
text(bp, 0, kmeans.clusters.table.structural, cex = 1, pos = 3)

kmeans.clusters.table.reduced <- table(kmeans.clusters.reduced$cluster)

bp <- barplot(kmeans.clusters.table.reduced, col = plot.colors, xlab = "clusters", ylab = "number of elements", 
              main = "Reduced data: k-means histrogram")
text(bp, 0, kmeans.clusters.table.reduced, cex = 1, pos = 3)
```

# Jaccard coefficient

```{r}
my.jaccard.coefficient <- function(labels1, cluster1, labels2, cluster2){
  labels1.comparisson <- labels1 == cluster1
  labels2.comparisson <- labels2 == cluster2
  return (sum(labels1.comparisson & labels2.comparisson ) / sum(labels1.comparisson | labels2.comparisson ))
}
```



```{r}
my.jaccard.matrix <- function(labels1, labels2, k){
  labels1 <- as.array(labels1)
  labels2 <- as.array(labels2)
  jaccard.hierarchical  <- matrix(nrow = k, ncol = k)
  max.coeffs.positions  <- array()
  max.coeffs            <- array()
  for (i in 1:k){
    max.coeffs.positions[i] <- 0
    max.coeffs[i] <- 0
    for(j in 1:k){
      jaccard.hierarchical[i,j] <- my.jaccard.coefficient(labels1, i, labels2, j)
      if (jaccard.hierarchical[i,j] > max.coeffs[i]){
        max.coeffs[i]           <- jaccard.hierarchical[i,j]
        max.coeffs.positions[i] <- j
      }
    }
  }
  return(list(jaccard.hierarchical, max.coeffs, max.coeffs.positions))
}
```



```{r}

jaccard.matrix.hierarchical <- my.jaccard.matrix(hierarch.clusters.cut.structural, hierarch.clusters.cut.reduced, k)
print(jaccard.matrix.hierarchical)
```


```{r}
jaccard.matrix.kmeans <- my.jaccard.matrix(kmeans.clusters.structural$cluster, kmeans.clusters.reduced$cluster, k)
print(jaccard.matrix.kmeans)
```
